---
title: "Abalone Project"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(here)
library(GGally)
library(readr)
library(CCA)
library(CCP)
library(MASS)
library(psych)
library(knitr)
library(dplyr)
library(e1071)
library(MVN)
library(purrr)
library(faraway)
```

```{r}

```

**Exploratory Data Analysis**

```{r}
abal <- read_csv(here("datasets","abalone.csv"))

```

```{r}
aba <- abal%>% mutate(Sex=factor(Sex))
```

```{r}
aba
```

```{r}
summary(aba)
```

```{r}
ggpairs(aba[,c(1:4)], aes(colour=Sex, alpha = 0.3))
```

In order to separate infants from the other abalone, We store the infant
abalone in dataset **iab**

```{r}
 iab <- aba %>% filter(Sex =="I") 
 iab 
```

The rest of the adult abalone in dataset **fmab**

```{r}
fmab <-aba %>% filter(Sex !="I")
fmab
```

We test for the difference in length, diameter, height of an infant
ablone and adult abalone with Welch two sample t- tests

```{r}
t.test(iab$Length,fmab$Length)
```

```{r}
t.test(iab$Diameter,fmab$Diameter)
```

```{r}
t.test(iab$Height, fmab$Height)
```

The p value of less than 2.2 x 10\^-16 in the three tests above shows
that the length, diameter, height of infant abalone is not similar to
adult abalone. These features can therefore be used to distinguish
between infants and adults.

A 95% confidence interval is used to correctly identify an infant
abalone from it Length Diameter and Height.

```{r}
t.test(iab$Length)
```

```{r}
t.test(iab$Diameter)
```

```{r}
t.test(iab$Height)
```

|              |          |                          |
|--------------|----------|--------------------------|
|              | **Mean** | **95% Confidence range** |
| **Length**   | 85.54 mm | 84.38 mm to 86.71 mm     |
| **Diameter** | 65.3 mm  | 64.35 mm to 66.24 mm     |
| **Height**   | 21.6 mm  | 21.25 mm to 21.94 mm     |

**Table 1. Range of Length , Diameter and Height of Infant abalone at
95% confidence**

From the above table any abalone which has a Length up to 87mm, Diameter
up to 67mm and Height up to 22mm can be considered as infants and
therefore should be avoided from harvesting.

For abalone beyond these dimensions we use a classification model which
classifies an abalone based on its length, diameter and height.

For this purpose we compare a Support Vector Machine model (SVM) ,
Linear Discriminant Analysis (LDA) and Quadratic Discriminant Analysis
(QDA)

**Testing for Multivariate Normality of Length, Diameter and Height**

```{r}
(mvn(aba[,c(2,3,4)]))$univariateNormality
```

As the three independent variables are not univairate or multivariate
normal, by default, SVM could the preferable classifier model

**Support Vector Machine (SVM)**

```{r}
set.seed(111)
ind <- sample(2, nrow(aba), 
              replace = TRUE, 
              prob = c(.8, .2))
training <- aba[ind==1,]
testing <- aba[ind==2,]
```

```{r}
(tuasvm <- tune.svm(Sex ~ Length+ Diameter + Height,data=training,kernel="radial", gamma = 10^(-1:1), cost = 10^(-1:1)))
```

```{r}
tuasvm$best.parameters
```

```{r}
tuasvm$best.model
```

```{r}
(asvm <- svm(Sex~Length + Diameter+ Height ,data = training, kernel="radial",gamma = tuasvm$best.parameters[1,1], cost = tuasvm$best.parameters[1,2], cross = 10))
```

```{r}
plot(asvm, data = training, Length~Diameter)
```

```{r}
plot(asvm, data = training, Length~Height)
```

```{r}
plot(asvm, data = training, Diameter~Height)
```

```{r}
pasvm <- predict(asvm, newdata = data.frame(Length = testing[,2], Diameter = testing[,3], Height = testing[,4]))
```

```{r}
(tasvm <-table(truth = testing$Sex, prediction = pasvm))
```

```{r}
sum(diag(tasvm)/sum(tasvm))*100
```

```{r}
tasvm[1.1]/nrow(testing %>% filter(Sex =="F"))*100
```

```{r}
tasvm[2,2]/nrow(testing %>% filter(Sex =="I"))*100
```

```{r}
tasvm[3,3]/nrow(testing %>% filter(Sex =="M"))*100
```

The SVM model has an accuracy of 50.7%

**Linear Discriminant Analysis (LDA)**

```{r}
summary(alda<- lda(Sex~Length + Diameter + Height, data = aba, CV = TRUE))
```

```{r}
(talda <- table(Truth = aba$Sex, Prediction = alda$class))
```

```{r}
sum(diag(talda)/sum(talda))*100
```

```{r}
talda[1,1]/nrow(aba %>% filter(Sex =="F"))*100
```

```{r}
talda[2,2]/nrow(aba %>% filter(Sex =="I"))*100
```

```{r}
talda[3,3]/nrow(aba %>% filter(Sex =="M"))*100
```

The LDA model has an accuracy of 51.88%

**Quadratic Discriminant analysis**

```{r}
summary(aqda<- qda(Sex~Length + Diameter + Height, data = aba, CV = TRUE))
```

```{r}
(taqda <- table(Truth = aba$Sex, Prediction = aqda$class))
```

```{r}
sum(diag(taqda)/sum(taqda))*100
```

```{r}
taqda[1,1]/nrow(aba %>% filter(Sex =="F"))*100
```

```{r}
talda[2,2]/nrow(aba %>% filter(Sex =="I"))*100
```

```{r}
talda[3,3]/nrow(aba %>% filter(Sex =="M"))*100
```

The QDA model has an accuracy of 52.15%

|                            |         |         |         |
|----------------------------|---------|---------|---------|
|                            | **SVM** | **LDA** | **QDA** |
| **Female**                 | 11.15%  | 22.34%  | 14.38%  |
| **Infant**                 | 74.89%  | 69.30%  | 69.29%  |
| **Male**                   | 64.68%  | 61.85%  | 61.84%  |
| **Overall**                | 50.76%  | 51.88%  | 52.15%  |
| **Misclassification Rate** | 49.24%  | 48.22%  | 47.85%  |

**Table 2. Classifier model accuracy**

**Testing for Multivariate Normality of Viscera Weight and Shucked
Weight**

```{r}
mvn(aba[,c(6,7)])
```

**Correlation Analysis**

```{r}
ggpairs(aba[,c(2,3,4,6,7)])
```

The two variables do not have univariate or multivariate normality

Correlation between the independent variables Length, Diameter and
Height

```{r}
cor(aba[,c(2,3,4)])
```

This shows shows that all the three independent variables are highly
correlated indicating high multicollinearity. A correlation between
length and diameter of 0.986 makes it possible for us to ignore one of
them in our modelling. We will test for its effect on our modelling. We
will however for now use all three Length, Diameter and Height.

Correlation between the dependent variables Viscera Weight, Shucked
Weight and Shell Weight

```{r}
cor(aba[,c(6,7,8)])
```

Checking for correlation between all, independent and dependent
variables

```{r}
cor(aba[,c(2,3,4,6,7,8)])
```

This shows shows that all the three dependent variables are also highly
correlated. Shell weight is of no commercial benefit and therefore can
be ignored in our modelling.

**Canonical Correlation Analysis**

```{r}
x<-aba[,c(2,3,4)]
y<-aba[,c(6,7)]
(abcor<-matcor(x,y))
```

```{r}
img.matcor(abcor, type = 2)
```

```{r}
summary(ccab <- cc(x,y))
```

```{r}
ccab$xcoef
```

```{r}
ccab$ycoef
```

```{r}
ccab$cor
```

```{r}
plt.cc(ccab, var.label = TRUE)
```

#### **Hypothesis Testing**

```{r}
n <- nrow(aba)
p <- ncol(x)
q <- ncol(y)
```

```{r}
p.asym(ccab$cor, n, p, q, tstat = "Wilks")
```

```{r}
p.asym(ccab$cor, n, p, q, tstat = "Pillai")
```

```{r}
p.asym(ccab$cor, n, p, q, tstat = "Hotelling")
```

The three hypothesis tests broadly agree and as the p value for both
canonical correlations is less than 0.05 we can consider that both
canonical correlations are statically significant.

However, as seen above are variables are not multivariate normal and the
population size is small so these test may generate some errors

**Permutation Testing**

```{r}
p.perm(x,y, type = "Wilks")
```

```{r}
p.perm(x,y, type = "Hotelling")
```

```{r}
p.perm(x,y, type = "Pillai")
```

The p value of 0 further confirms that the first canonical correlations
are statistically significant.

**Multivariate Linear Modeling and Analysis**

We would build a linear model to predict shucked weight and visceral
weight by regressing over independent variables of Length, Diameter and
Height.

```{r}
amlm <- lm(cbind(`Shucked weight`,`Viscera weight`)~Length+Diameter+Height, data=aba)
summary(amlm)
```

```{r}
coef(amlm)
```

```{r}
confint(amlm, level = 0.95)
```

```{r}
vcov(amlm)
```

```{r}
head(resid(amlm))
```

**Hypothesis Testing by ANOVA**

```{r}
anova(amlm, test="Wilks")
```

As we have seen above the correlation between Length and Diameter is
0.98 making them highly correlated. We therefore test if diameter has an
extra predictive effect on the shucked and viscera weights.

```{r}
ramlm<- lm(cbind(`Shucked weight`,`Viscera weight`)~Length+Height, data=aba)
anova(amlm,ramlm)
```

This test gives a p value less than 0.05 and therefore conclude that
Diameter has an effect on the prediction of shucked and viscera weight

```{r}
predict(amlm, newdata = testing[1:5,] ,interval ='confidence')
```

**Pearson Residuals**

```{r}
sds <- amlm %>% estVar %>% diag %>% sqrt
PR <- sweep(resid(amlm), 2, sds,)
ggpairs(as_tibble(PR))
```

```{r}
res<- resid(amlm)
plot(fitted(amlm), res)
abline(0,0)
```

The above two graphs of residues show that this linear model is good fit
for the given data.

We can therefore conclude that we have a linear model for each, shucked
and viscera weight

**Viscera Weight = -48.339 + 0.502 Length + 0.255 Diameter + 0.391
Height**

**Shucked Weight = -99.107 + 1.352 Length + 0.474 Diameter + 0.476
Height**

Please access the original RMD file from

<https://drive.google.com/file/d/1jneyhTl8bTup8qs_N19h3H0BOcQbzBzd/view?usp=share_link>
